Filamentary transport in high-power H-mode conditions and in no/small-ELM regimes to predict heat and particle loads on PFCs for future devices

### 2016-11-09 Presentation/gpm_november2016_vianello.tex
Added a bunch of figures and update to slides
Finished the L-Mode description from AUG/TCV/JET

\usepackage[style=nature,citestyle=authoryear-comp,defernumbers=true,maxnames=2,firstinits=true,
uniquename=init,backend=bibtex8,arxiv=abs,mcite]{biblatex}
\bibliography{biblio}
\renewcommand*{\bibfont}{\footnotesize}
\renewcommand*{\citesetup}{\footnotesize}
\usepackage[export]{adjustbox}
\makeatother
\mode<presentation>
\makeatletter
% add a macro that saves its argument
\newcommand{\footlineextra}[1]{\gdef\insertfootlineextra{#1}}
\newbox\footlineextrabox
% for reducing font on a single slide
\newcommand\Fontvi{\fontsize{8}{7.2}\selectfont}
\title{Topic 21: Filamentary transport in high-power H-mode conditions and in
  no/small-ELM regimes to predict heat and particle loads on PFCs for
  future devices}
\date{15 November 2016}
\author[N.Vianello]{J. Madsen and N. Vianello}
\begin{document}
\tikzstyle{every picture}+=[remember picture]
\maketitle
\begin{frame}{Present status}
  \setbeamercovered{transparent}
  \begin{enumerate}[<+(1) | invisible@-+>]
    \item Recent experimental and theoretical investigation agrees on
      suggesting SOL density profiles is determined by a balance
      between outward convective filamentary transport and parallel dynamics
    \item The

  \end{itemize}
\end{frame}
% \begin{frame}{Background and objectives}
% \only<1>{
% \begin{itemize}
% \item Role of turbulence transport in the SOL saturation is a well
%   know
%   feature \parencite{LaBombard:2001ks,Rudakov:2005ic,Garcia:2007hh,Carralero:2014gs}. \textcolor{red}{Increasing
%   density, even without reaching detachment SOL profile tend to flatten}
% \end{itemize}}
% \only<2-3>{
%   \begin{columns}
%     \begin{column}{0.5\textwidth}
%       \begin{itemize}
%       \item Present understanding suggests that a transition of filament
%         propagation regime is dominated by effective
%         collisionality $\Lambda$ \parencite{Myra:2006p2754}
%         \begin{equation*}
%           \Lambda = \frac{L_{\parallel}/c_s}{1/\nu_{ei}}\frac{\Omega_i}{\Omega_e}
%         \end{equation*}
%         \item<3> \textcolor{red}{At constant blob size, increasing
%             collisionality provide a change in the filaments velocity
%             scaling properties}
%       \end{itemize}
%     \end{column}
%     \begin{column}{0.5\textwidth}
% %      \includegraphics<2>[width=\textwidth]{../pdfbox/Presentation/FigMyra}
%         \begin{tikzonimage}[width=\textwidth]{../pdfbox/Presentation/FigMyra}
%           \only<3>{
%             \draw [->, ultra thick, red] (0.7,0.3) -- (0.7,0.6);}
%         \end{tikzonimage}
%     \end{column}
%   \end{columns}
% }
% \only<4-6>{
%   \begin{columns}
%     \begin{column}{0.5\textwidth}
%       \begin{itemize}
%       \item AUG and JET \parencite{Carralero:2015gu} suggest that $\Lambda_{div}$ dominates this
%         process and a transition from
%         \textcolor{ta3chameleon}{sheath-limited}
%         \onslide<5-6>{to \textcolor{blue}{inertial regime}} 
%         \item<6> \textcolor{red}{Tested by changing n$_e$
%             and $T_e$ through fueling/seeding/heating}
%       \end{itemize}
%     \end{column}
%     \begin{column}{0.5\textwidth}
%       \begin{tikzonimage}[width=\textwidth]{../pdfbox/Presentation/KoMFig2.png}
%         \only<4-6>{
%           \draw [thick, ta3chameleon, thick] (0.45,0.25) ellipse (0.23 and 0.1);
%       }
%         \only<5-6>{
%           \draw [thick, blue, thick, rotate=-4] (0.8,0.55) ellipse (0.1 and 0.33);
%       }
%         \only<6>{
%           \draw [thick, red, thick, dashed] (0.35,0.75) circle (0.25);
%       }
%     \end{tikzonimage}
%   \end{column}
%   \end{columns}
% }
% \only<7->{
%   \begin{columns}
%     \begin{column}{0.5\textwidth}
%       \begin{itemize}
%       \item MAST \parencite{Militello:2016hk} and old TCV
%         \parencite{Garcia:2007p2615} data 
%         suggested and I$_P$ dependence. 
%           \textcolor{red}{At higher current and
%         same density SOL shoulder disappear} 
%       \item<8> \textcolor{ta3chameleon}{Being shoulder formation
%           induced by a change of $\Gamma_{\perp}$ \textit{vs}
%           $\Gamma_{\parallel}$ balance we will try to test dependence
%           on L$_{\parallel}$}
%     \end{itemize}
%     \end{column}
%     \begin{column}{0.5\textwidth}
%       \includegraphics[width=\textwidth]{../pdfbox/Presentation/FigMilitello.pdf}
%     \end{column}
%   \end{columns}
% }
% \centering{\includegraphics<1>[width=1.1\textwidth]{../pdfbox/Presentation/Fig1Presentation}}
% \end{frame}

>>>>

\usepackage[style=nature,citestyle=authoryear-comp,defernumbers=true,maxnames=2,firstinits=true,
uniquename=init,backend=bibtex8,arxiv=abs,mcite]{biblatex}
\bibliography{biblio}
\renewcommand*{\bibfont}{\footnotesize}
\renewcommand*{\citesetup}{\footnotesize}
\usepackage[export]{adjustbox}
\makeatother
\mode<presentation>
\makeatletter
% add a macro that saves its argument
\newcommand{\footlineextra}[1]{\gdef\insertfootlineextra{#1}}
\newbox\footlineextrabox
% for reducing font on a single slide
\newcommand\Fontvi{\fontsize{8}{7.2}\selectfont}
\title{Filamentary transport in high-power H-mode conditions and in
  no/small-ELM regimes to predict heat and particle loads on PFCs for
  future devices}
\date{15 November 2016}
\author[J. Madsen and N.Vianello]{J. Madsen and N. Vianello}
\begin{document}
\tikzstyle{every picture}+=[remember picture]
\maketitle
\begin{frame}{Present status}
  \setbeamercovered{transparent}
  \begin{enumerate}[<+(1) | invisible@-+>]
    \item Recent experimental and theoretical investigation agrees on
      suggesting SOL density profiles is determined by a balance
      between radial convective filamentary transport and parallel dynamics
    \item The balance is observed to change at high density in L-Mode
      and this is at the basis of the \emph{shoulder formation}
    \item This is observed in a variety of devices including AUG \parencite{Carralero:2014gs},
      TCV \parencite{Garcia:2007p2615} and MAST \parencite{Militello:2016hk}
    \item The observations exhibit various dependence on Divertor
      Collisionality, Plasma current,  seeding. \alert{Results from
        different devices not fully reconciled.}
    \item The H-Mode behavior is still a work in progress and exhibit
      a dependence on the level of \alert{fueling}
    \item First we will report a summary of the achievements and then
     personal comments and plans
    \end{enumerate}
\end{frame}
\section{L-Mode}
\begin{frame}{L-Mode studies: AUG/1}
  \begin{columns}
    \begin{column}{0.5\textwidth}
      \begin{itemize}
      \item<1-> AUG and JET \parencite{Carralero:2015gu} suggest that
        \begin{equation*}
        \Lambda_{div} = \frac{L_{\parallel}/c_s}{1/\nu_{ei}}\frac{\Omega_i}{\Omega_e}
      \end{equation*}
      dominates this
        process and a transition from
        \textcolor{ta3chameleon}{sheath-limited}
        to \textcolor{blue}{inertial regime}
        \item<2-> \textcolor{red}{Tested by changing n$_e$
            and $T_e$ through fueling/seeding/heating}
        \item<3-> This determines a change of the velocity-size
          scaling. \onslide<4>{\alert{$\Lambda_{div}$ rules the density profile scale
          length}} 
        \end{itemize}
    \end{column}
    \begin{column}{0.5\textwidth}
      \only<1-2>{
      \begin{tikzonimage}[width=\textwidth]{../pdfbox/KoM15Nov/CarraleroPRL15a}
          \draw [thick, ta3chameleon, thick, rotate=1] (0.43,0.25) ellipse (0.23 and 0.1);
          \draw [thick, blue, thick, rotate=-4] (0.8,0.55) ellipse (0.1 and 0.33);
        \only<2>{
          \draw [ultra thick, red, thick, dashed] (0.5,0.63) circle (0.12);
        }
      \end{tikzonimage}}
    \only<3>{
      \begin{tikzonimage}[width=\textwidth]{../pdfbox/KoM15Nov/CarraleroPRL15b}
          \draw [ultra thick, ta3chameleon, rotate=1] (0.2,0.5) ellipse (0.05 and 0.4);
          \draw [thick, blue, rotate=15] (0.65,0.07) ellipse (0.4 and 0.1);
      \end{tikzonimage}
    }
    \includegraphics<4>[width=\textwidth]{../pdfbox/KoM15Nov/CarraleroPRL15c}
    \end{column}
  \end{columns}
\end{frame}
\begin{frame}{L-Mode studies: AUG/2}
  \begin{columns}
    \begin{column}{0.5\textwidth}
      \begin{itemize}
      \item<1-> Profile modified by increase of blob-size and change of
        \alert{packing fraction: $f_{fil} = \nu_{fil}\tau_{AC}$}
        \onslide<2->{\textcolor{blue}{and filament relative density}
          {\footnotesize (Carralero 2016 in preparation)}}
      \item<3-> As a consequence the contribution of filaments to
        radial transport increases
      \item<4> Parallel flow is strongly reduced whenever we increase
        the divertor collisionality
      \end{itemize}
    \end{column}
      \begin{column}{0.5\textwidth}
        \includegraphics<1>[width=\textwidth]{../pdfbox/KoM15Nov/CarraleroMST16a}
        \includegraphics<2>[width=\textwidth]{../pdfbox/KoM15Nov/CarraleroMST16b}
        \includegraphics<3>[width=\textwidth]{../pdfbox/KoM15Nov/CarraleroMST16c}
        \includegraphics<4>[width=\textwidth]{../pdfbox/KoM15Nov/CarraleroMST16d}
      \end{column}
    \end{columns}
  \end{frame}

  \begin{frame}{L-Mode studies:AUG/3}
  \begin{columns}
    \begin{column}{0.5\textwidth}
      \begin{itemize}
      \item<1-> Electron and ions behave differently
      \item<2|only@2> T$_{e, fil} \sim 1.2 $T$_{e, bk}$ roughly constant
        accross the SOL and slightly affected by the increase of
        divertor collisionality
      \item<3|only@3> Ions are strongly affected: for
        \textcolor{red}{$\Lambda_{div}<1$ T$_{i, fil} > $ T$_{i, bk}$
          and $\lambda_{T_i} \sim 30$
          mm}. \textcolor{blue}{$\Lambda_{div}>1$ T$_{i, fil} \sim $
          T$_{i, bk}\sim 25$ eV
          and $\lambda_{T_i} \sim 8$ mm}
       \item<4-> Ion energy spectrum from
         $\mathbf{E}\times\mathbf{B}$ analyzer shrinks towards lower
         energy for $\Lambda_{div} > 1$
       \item<5> EMC3-Eirene simulation suggests that such a reduction
         can't be accounted for thermalization process. An
         \alert{ionization front builds in front of the limiter shadow}   
       \end{itemize}
    \end{column}
      \begin{column}{0.5\textwidth}
        \includegraphics<2>[width=\textwidth]{../pdfbox/KoM15Nov/CarraleroMST16e}
        \includegraphics<3>[width=\textwidth]{../pdfbox/KoM15Nov/CarraleroMST16f}
        \includegraphics<4>[width=\textwidth]{../pdfbox/KoM15Nov/CarraleroMST16g}
        \includegraphics<5>[width=\textwidth]{../pdfbox/KoM15Nov/CarraleroMST16h}
      \end{column}
    \end{columns}
  \end{frame}

  \begin{frame}{L-Mode: TCV}
    \begin{columns}
    \begin{column}{0.5\textwidth}
      \begin{itemize}
      \item<1|only@1> Flexibility has allowed to test $\Lambda_{div}$
        dependence on L$_{\parallel}$ by varying flux expansion f$_x$:
        \begin{equation*}
          f_x = \frac{(B_p/B_t)_{MP}}{(B_p/B_t)_{SP}}
        \end{equation*}
        in ohmic density ramps \parencite{vianello:iaea2016}
      \item<2|only@2> Slight variation of density profiles at the target but
        due to direct dependence on L$_{\parallel}$ large increase of
        $\Lambda_{div}$. \alert{Upstream profiles only varies whenever
        we reach a certain amount of $\langle n_e \rangle$}
       \only<3-5>{\item Weak dependence of blob-size from $\Lambda_{div}$,
        \onslide<4-5>{\textcolor{red}{also on a statistical
            basis}}. \onslide<5>{Strong dependence on average density,
          independent of L$_{\parallel}$}}
       \item<6|only@6> $\lambda_n$ depends clearly on blob-size
         whereas the dependence on divertor condition is less
         obvious. \alert{$\Lambda_{div}$ necessary but not sufficient}
      \end{itemize}
    \end{column}
      \begin{column}{0.5\textwidth}
        \includegraphics<2>[width=\textwidth]{../pdfbox/KoM15Nov/VianelloIAEA16a}
        \includegraphics<3>[width=\textwidth]{../pdfbox/KoM15Nov/VianelloIAEA16b}
        \includegraphics<4>[width=\textwidth]{../pdfbox/KoM15Nov/VianelloIAEA16c}
        \includegraphics<5>[width=\textwidth]{../pdfbox/KoM15Nov/VianelloIAEA16d}
        \includegraphics<6>[width=\textwidth]{../pdfbox/KoM15Nov/VianelloIAEA16e}
      \end{column}
    \end{columns}
  \end{frame}

  \begin{frame}{L-Mode: JET}
    \begin{columns}
    \begin{column}{0.5\textwidth}
      \begin{itemize}
      \item<1|only@1> The shoulder formation strongly depends on
        divertor geometry, disappear with vertical target and strike
        point closest to cryogenics pumps \parencite{Wynn:EPS2016}
      \item<2|only@2> The midplane pressure from baratrons is
        equivalent between the different divertor. \alert{This would
          indicate that SOL neutral density does not play any role}
       \item<3|only@3> In the horizontal target configuration the
         results indicate that the shoulder forms right at the
         transition from sheath-limited to high-recycling where also
         $\Lambda_{div}$ strongly increase
       \item<4|only@4> Shoulder amplitude correlates with strike
         points position. \alert{Shoulder, ionization and
           $\Gamma_{ion, plate}$ larger when R$_{strike}$ smaller away
         from the pump}
       \item<5|only@5> In seeded discharges the transition observed at
         very high level of $\Lambda_{div} >> 1$
      \end{itemize}
    \end{column}
      \begin{column}{0.5\textwidth}
        \includegraphics<1>[width=\textwidth]{../pdfbox/KoM15Nov/LipschultzITPA16a}
        \includegraphics<2>[width=\textwidth]{../pdfbox/KoM15Nov/LipschultzITPA16b}
        \includegraphics<3>[width=\textwidth]{../pdfbox/KoM15Nov/LipschultzITPA16c}
        \includegraphics<4>[width=\textwidth]{../pdfbox/KoM15Nov/LipschultzITPA16d}
        \includegraphics<5>[width=\textwidth]{../pdfbox/KoM15Nov/LipschultzITPA16e}
      \end{column}
    \end{columns}

  \end{frame}
  % \only<1>{
% \begin{itemize}
% \item Role of turbulence transport in the SOL saturation is a well
%   know
%   feature \parencite{LaBombard:2001ks,Rudakov:2005ic,Garcia:2007hh,Carralero:2014gs}. \textcolor{red}{Increasing
%   density, even without reaching detachment SOL profile tend to flatten}
% \end{itemize}}
% \only<2-3>{
%   \begin{columns}
%     \begin{column}{0.5\textwidth}
%       \begin{itemize}
%       \item Present understanding suggests that a transition of filament
%         propagation regime is dominated by effective
%         collisionality $\Lambda$ \parencite{Myra:2006p2754}
%         \begin{equation*}
%           \Lambda = \frac{L_{\parallel}/c_s}{1/\nu_{ei}}\frac{\Omega_i}{\Omega_e}
%         \end{equation*}
%         \item<3> \textcolor{red}{At constant blob size, increasing
%             collisionality provide a change in the filaments velocity
%             scaling properties}
%       \end{itemize}
%     \end{column}
%     \begin{column}{0.5\textwidth}
% %      \includegraphics<2>[width=\textwidth]{../pdfbox/Presentation/FigMyra}
%         \begin{tikzonimage}[width=\textwidth]{../pdfbox/Presentation/FigMyra}
%           \only<3>{
%             \draw [->, ultra thick, red] (0.7,0.3) -- (0.7,0.6);}
%         \end{tikzonimage}
%     \end{column}
%   \end{columns}
% }
% \only<4-6>{
%   \begin{columns}
%     \begin{column}{0.5\textwidth}
%       \begin{itemize}
%       \item AUG and JET \parencite{Carralero:2015gu} suggest that $\Lambda_{div}$ dominates this
%         process and a transition from
%         \textcolor{ta3chameleon}{sheath-limited}
%         \onslide<5-6>{to \textcolor{blue}{inertial regime}} 
%         \item<6> \textcolor{red}{Tested by changing n$_e$
%             and $T_e$ through fueling/seeding/heating}
%       \end{itemize}
%     \end{column}
%     \begin{column}{0.5\textwidth}
%       \begin{tikzonimage}[width=\textwidth]{../pdfbox/Presentation/KoMFig2.png}
%         \only<4-6>{
%           \draw [thick, ta3chameleon, thick] (0.45,0.25) ellipse (0.23 and 0.1);
%       }
%         \only<5-6>{
%           \draw [thick, blue, thick, rotate=-4] (0.8,0.55) ellipse (0.1 and 0.33);
%       }
%         \only<6>{
%           \draw [thick, red, thick, dashed] (0.35,0.75) circle (0.25);
%       }
%     \end{tikzonimage}
%   \end{column}
%   \end{columns}
% }
% \only<7->{
%   \begin{columns}
%     \begin{column}{0.5\textwidth}
%       \begin{itemize}
%       \item MAST \parencite{Militello:2016hk} and old TCV
%         \parencite{Garcia:2007p2615} data 
%         suggested and I$_P$ dependence. 
%           \textcolor{red}{At higher current and
%         same density SOL shoulder disappear} 
%       \item<8> \textcolor{ta3chameleon}{Being shoulder formation
%           induced by a change of $\Gamma_{\perp}$ \textit{vs}
%           $\Gamma_{\parallel}$ balance we will try to test dependence
%           on L$_{\parallel}$}
%     \end{itemize}
%     \end{column}
%     \begin{column}{0.5\textwidth}
%       \includegraphics[width=\textwidth]{../pdfbox/Presentation/FigMilitello.pdf}
%     \end{column}
%   \end{columns}
% }
% \centering{\includegraphics<1>[width=1.1\textwidth]{../pdfbox/Presentation/Fig1Presentation}}
% \end{frame}
